### 2021-08-12 paper/sections/numerical.tex
Set up the repo
- Add all functions and scripts necessary for the merger initialisation pipeline

- Improved parameter file design

\subsection{The KETJU Code}
\drafting{Basic overview and updates, refer to previous papers}

\subsection{Initial Conditions and Simulations}