### 2021-08-12 paper/sections/results.tex

\label{sec:results}

\drafting{
Figures to do
\begin{itemize}
\item Good ones
\end{itemize}
Table of galaxy properties,
and maybe another one of BH merger remnant properties.
}

%\begin{figure*}
%\includegraphics{binary_params_main}
%\includegraphics{binary_params_secondary}
%\includegraphics{binary_params_tertiary}
%\caption{
%\drafting{
%Binary orbital parameters. Dashed line for triplet outer orbit (dashes not very %visible at the moment).
%Lower panels are mergers occurring in separate galaxies.
%Note the use of $1-e$ for the last panel.
%}
%}
%\label{fig:binary_params}
%\end{figure*}