### 2022-07-26 paper/sections/results.tex
Small updates to paper draft

\label{sec:results}

\drafting{
Figures to do
\begin{itemize}
\item Example of forward-modelled semimajor axis and eccentricity
\item Example of forward-modelled density profile
\item Example of forward-modelled velocity dispersion profile
\end{itemize}
Table of galaxy properties,
and maybe another one of BH merger remnant properties.
}

\drafting{
    WHAT ARE THE MAIN OBSERVATIONS WE SEE and WHAT CONCLUSIONS CAN WE DRAW?
    \begin{enumerate}
        \item stability of hardening rate -> semimajor axis can be robustly predicted
        \item wide variance of eccentricity -> N-body models offer no quantitative prediction on likelihood of merger, can only comment on what merger looks like \textit{IF} it happens
        \item variance in observables: trends between observable properties and what observers can actually make use of (don't have this yet), e.g. signatures of SMBH scouring in velocity profile?
    \end{enumerate}
}

%\begin{figure*}
%\includegraphics{binary_params_main}
%\includegraphics{binary_params_secondary}
%\includegraphics{binary_params_tertiary}
%\caption{
%\drafting{
%Binary orbital parameters. Dashed line for triplet outer orbit (dashes not very %visible at the moment).
%Lower panels are mergers occurring in separate galaxies.
%Note the use of $1-e$ for the last panel.
%}
%}
%\label{fig:binary_params}
%\end{figure*}