### 2022-08-26 paper/sections/stats.tex

We perform full Bayesian Hierarchical Modelling for the projected mass density profiles of our simulated galaxies. 
The modelling is performed using the statistical software \textsc{Stan}\footnote{https://mc-stan.org}, which itself implements Hamiltonian Monte Carlo (HMC).

\subsection{Hamiltonian Monte Carlo}
An excellent overview of HMC may be found in \citet{betancourt17}, with a brief description provided here.

In numerically sampling an arbitrary target distribution, we are required to sample from the \textit{typical set}, or those regions of the target distribution which provide the dominant contribution to an expectation value. 
Simplistically, this may be achieved using any Markov generating process, such as the widely-used Metropolis-Hastings (MH) algorithm, albeit at the expense of inefficient and biased sampling. 
The MH algorithm generates from a starting position $q$ a proposal for the next position $q'$, with proposals accepted with some probability $Q(q'|q)$, where $Q$ is often chosen to be the normal distribution. 
Sampling with increased efficiency can be achieved by using information about the geometry of the typical set of the target distribution through a vector field aligned with the typical set, which may be estimated using the gradient of the target distribution. 
To align the gradient vector field with the typical set vector field, we can introduce auxillary momenta variables $p$, promoting our $D$-dimensional target distribution to a projection of a $2D$-dimensional canonical distribution in some phase space:
\begin{equation*}
    P(q,p) = P(p|q)P(q) = e^{-H(q,p)}
\end{equation*}
where $H$ is an invariant Hamiltonian function. 
The desired vector field may then be generated from the Hamiltonian $H$ using Hamilton's equations. 
This formulation also lends itself to improved sampling efficiency by decomposing the Hamiltonian into a \textit{kinetic energy} term and a \textit{potential energy} term, with the latter equivalent to the logarithm of the target distribution. 
As the Hamiltonian is phase space conserving, they may be accurately integrated using a volume-preserving integration scheme, such as symplectic integrators. 
Of these, the leapfrog scheme offers both a simple implementation and high accuracy over long integration times.

\subsection{Hierarchical Models}
\begin{itemize}
    \item Data doesn't have error, model has uncertainty
    \item Require multiple observations from a population --> samples
    \item Assume population described by a series of hyperparameters
    \item Prior distributions on those hyperparameters
    \item Forward fold through model, combine with likelihood function from data
\end{itemize}

\drafting{\begin{enumerate}
    \item HMC algorithm?
    \item Description of hierarchical model
    \item Prior distributions
    \item forward modelling
\end{enumerate}
}