### 2022-09-21 papers/paper-1/sections/intro.tex

\section{Introduction}

\drafting{Something cool}

\drafting{
    Introduce core sersic profile here.
}
\begin{equation}\label{eq:coresersic}
    \Sigma(R) = \Sigma'\left[ 1+\left(\frac{r_\mathrm{b}}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[ -b_n \left(\frac{R^\alpha+r_\mathrm{b}^\alpha}{R_\mathrm{e}^\alpha}\right)^{1/(\alpha n)} \right],
\end{equation}
where
\begin{equation}\label{eq:coresersic2}
    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[ b_n \left( 2^{1/\alpha} r_\mathrm{b}/R_\mathrm{e} \right)\right]
\end{equation}