### 2022-09-21 papers/paper-1/sections/numerical.tex

\section{Numerical Simulations}

\drafting{Describe the numerical set up, code, etc}