### 2022-09-21 papers/paper-2/sections/results.tex

\section{Results}

\drafting{Have some nice plots, groundbreaking results}

\begin{figure*}
    \includegraphics{compare_AC_gqs.png}
    \caption{\drafting{Comparing the latent parameters for different members of the same progenitor class. Here shown is those families part of the (approximately equal-mass) AC merger class. There is no significant difference in latent parameter values with increasing rperi. TODO do this same plot for each merger class.}}
    \label{fig:compareAC}
\end{figure*}

\begin{figure*}
    \includegraphics{compare_AD_gqs.pdf}
    \caption{\drafting{Same as \autoref{fig:compareAC}, but for the (unequal mass) AD merger class. For merger families with $R_\mathrm{peri}/R_{\mathrm{vir},1}\leq0.1$, the latent parameter distributions are indistinguishable. COnversely, for the A-D-3.0-1.0 family, the effective radius has its mode at lower values than the other merger class members, and its S\'ersic index peaks at slightly higher values of $n$.}}
    \label{fig:compareAD}
\end{figure*}