### 2023-02-21 papers/paper-1/diagrams/hm_coresersic.tex
Small updates in preparation for SPZ meeting

\documentclass[border=10pt]{standalone}
\usepackage{newtxtext, newtxmath}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\begin{document}
\begin{tikzpicture}

    % define nodes
    % define layer 1
    \node[obs] (Shat) {$\hat{\Sigma}$};
    \factor[below=of Shat] {S_f} {right:Lognorm} {} {};
    \node[det, below=of S_f] (S) {$\Sigma$};
    \node[obs, below=of S] (R) {$\hat{R}$};
    \node[obs, right=of S] (sigma) {$\hat{\sigma_\Sigma}^2$};

    % define layer 2
    \node[latent, left=of S] (Sb) {$\tilde{\Sigma}_\mathrm{b}$};
    \node[latent, below=of Sb] (Re) {$R_\mathrm{e}$};
    \node[latent, below=of Re] (rb) {$r_\mathrm{b}$};
    \node[latent, above=of Sb] (g) {$\gamma$};
    \node[latent, above=of g] (n) {$n$};
    \node[latent, above=of n] (a) {$a$};

    % define layer 3
    \factor[left=of rb] {rb_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of rb_f, yshift=-1.2em] (rbs) {$r_{\mathrm{b},\sigma}$};
    \node[hyper, left=of rb_f, yshift=1.2em] (rbm) {$r_{\mathrm{b},\mu}$};

    \factor[left=of Re] {Re_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of Re_f, yshift=-1.2em] (Res) {$R_{\mathrm{e},\sigma}$};
    \node[hyper, left=of Re_f, yshift=1.2em] (Rem) {$R_{\mathrm{e},\mu}$};

    \factor[left=of Sb] {Sb_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of Sb_f, yshift=-1.2em] (Sbs) {$\tilde{\Sigma}_{\mathrm{b},\sigma}$};
    \node[hyper, left=of Sb_f, yshift=1.2em] (Sbm) {$\tilde{\Sigma}_{\mathrm{b},\mu}$};

    \factor[left=of g] {g_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of g_f, yshift=-1.2em] (gs) {$\gamma_\sigma$};
    \node[hyper, left=of g_f, yshift=1.2em] (gm) {$\gamma_\mu$};

    \factor[left=of n] {n_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of n_f, yshift=-1.2em] (ns) {$n_\sigma$};
    \node[hyper, left=of n_f, yshift=1.2em] (nm) {$n_\mu$};

    \factor[left=of a] {a_f} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of a_f, yshift=-1.2em] (as) {$a_\sigma$};
    \node[hyper, left=of a_f, yshift=1.2em] (am) {$a_\mu$};

    % connect the nodes
    \factoredge {sigma, S} {S_f} {Shat};

    \edge {rb, Re, Sb, g, n, a, R} {S};

    \factoredge {rbs, rbm} {rb_f} {rb};
    \factoredge {Res, Rem} {Re_f} {Re};
    \factoredge {Sbs, Sbm} {Sb_f} {Sb};
    \factoredge {gs, gm} {g_f} {g};
    \factoredge {ns, nm} {n_f} {n};
    \factoredge {as, am} {a_f} {a};

    % plates
    \plate {rS} {(R)(S)(sigma)(Shat)} {$\forall R,\sigma_\Sigma^2$};
    \plate {pars} {(R)(S)(sigma)(rb)(Re)(Sb)(g)(n)(a)} {$\forall$ child};

\end{tikzpicture}
\end{document}