### 2023-03-11 papers/paper-1/sections/appendix.tex
Start to edit paper 1 to reflect project direction change

\section{Prior Predictive Checks for Projected Mass Density}
Here we show the prior predictive check for the projected mass density of the H-1.00 system, as well the corner plots of the latent parameter hyperdistributions.