### 2023-03-11 papers/paper-1/sections/numerical.tex

\section{Numerical Simulations}

\subsection{Initial Conditions}

\begin{table*}
    \centering
    \caption{Parameters of Hernquist sphere galaxy models used in determining the binary hardening convergence.}
    \label{tab:hernquist_pars}
    \begin{tabular}{cccccccc}
        \hline
        Name & $m_\star / (10^5 \Msun)$ & $N_\star/10^6$ & $\varepsilon_\star$/pc & $m_\mathrm{DM} / (10^7 \Msun)$ & $N_\mathrm{DM}/10^6$ & $\varepsilon_\mathrm{DM}$/pc & $M_\bullet/m_\star$\\
        \hline
        H-1.000 & 0.5 & 5.02 & 5.0 & 1.0 & 4.54 & 300.0 & $6.0\times 10^4$ \\
        H-0.500 & 1.0 & 2.51 & 5.0 & 2.0 & 2.27 & 300.0 & $3.0\times 10^4$ \\
        H-0.250 & 2.0 & 1.25 & 5.0 & 4.0 & 1.13 & 300.0 & $1.5\times 10^4$ \\
        H-0.100 & 5.0 & 0.50 & 5.0 & 10.0 & 0.45 & 300.0 & $6.0\times 10^3$ \\
        H-0.050 & 10.0 & 0.25 & 10.0 & 20.0 & 0.23 & 600.0 & $3.0\times 10^3$ \\
        H-0.010 & 50.0 & 0.05 & 10.0 & 100.0 & 0.04 & 600.0 & $6.0\times 10^2$ \\
        H-0.005 & 100.0 & 0.025 & 20.0 & 200.0 & 0.02 & 1000.0 & $3.0\times 10^1$ \\
        H-0.001 & 500.0 & 0.005 & 20.0 & 1000.0 & 0.004 & 1000.0 & $6.0\times 10^1$ \\
        \hline
    \end{tabular}
\end{table*}

We simulate the merger of two identical galaxies. 
Each galaxy is modelled as a spherically-symmetric multicomponent system: a stellar component embedded within a DM halo, and a single SMBH at the centre of the galaxy. 
The total stellar mass is set to $M_\star=2.5\times10^{11}\,\Msun$, the total DM mass is set to $M_\mathrm{DM} \sim 4.9\times 10^{13}\,\Msun$, and the SMBH mass is set to $M_\bullet \sim 3.0\times10^9\,\Msun$.
The total DM mass and BH mass are determined as a function of total stellar mass from the relations of \citet{behroozi19} and \citet{sahu19}, respectively. 
Both the stellar and DM components have a density profile given by an ergodic Hernquist sphere parametrised by scale radius $a$:
\begin{equation}\label{eq:hernquist}
    \rho(r) = \frac{2M}{4\pi} \frac{a}{r(r+a)^3}.
\end{equation}
We use a value of $a_\star=3.0\,\mathrm{kpc}$ for the stellar component, and a value of $a_\mathrm{DM}=300.0\,\mathrm{kpc}$ for the DM component. 
We Monte Carlo sample the phase space distribution function
\begin{equation}\label{eq:eddington}
    f(\mathcal{E}) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^\mathcal{E} \dv{\Psi}{\sqrt{\mathcal{E}-\Psi}} \dv{^2\nu_\mathcal{E}}{\Psi^2} + \frac{1}{\sqrt{\mathcal{E}}} \left(\dv{\nu_\mathcal{E}}{\Psi}\right)_{\Psi=0}\right],
\end{equation}
to generate position and velocity coordinates for all stellar and DM particles.
As mentioned, the SMBH is initialised at rest at the origin.
Additionally, we give the SMBH a randomly-drawn value of the dimensionless spin parameter $\chi$ following \drafting{one of the Zlochower papers}, with shape parameters $\alpha=10.5868, \beta=4.66884$. 
This distribution has a mode at $\chi \sim 0.8$. 
We test eight mass resolutions between the SMBH mass and the stellar particle mass, with four realisations generated for each mass resolution by independently Monte Carlo sampling the phase space distribution function.

We simulate the equal-mass merger between like-mass resolution systems for each mass resolution listed in \autoref{tab:hernquist_pars} for all combination of galaxy pairs (i.e., ten mergers per mass resolution).
The progenitor galaxies in each merger are set to an initial separation of $20\,\mathrm{kpc}$ on quasi-Keplerian orbit, assuming each galaxy is a point-mass containing half its total mass. 
The initial orbital eccentricity is chosen to be almost radial ($e_0 = 0.99$), and the first pericentre distance $r_\mathrm{peri}$ is set to $\sim 8\,\mathrm{kpc}$.

\subsection{Simulations}

The progenitors are evolved with \ketju{}.
Within a relatively short time frame, the SMBH binary hardens predominantly through dynamical friction and three-body stellar interactions. 
We evolve the binary to the time $t_{a_h}=t(a_h)$, and continue to at least when $t_\text{safe}=t(a=10\,\text{pc})$. 
As we do not require the gravitational wave dominated regime to be reached for our analysis, we may terminate the runs before the SMBH binary merges.