### 2023-04-19 papers/paper-eccentricity/sections/numerical.tex
Set up directory for eccentricity paper

\section{Numerical Simulations}

\begin{itemize}
    \item describe toy model
    \item describe galaxy ICs: 0.5 Dehnen spheres
    \item describe merger orbit: low and high e
\end{itemize}