### 2023-05-09 papers/paper-1/sections/intro.tex

\section{Introduction}

\drafting{There was a paper by \citet{nasim20} about ``defeating stochasticity''. They use some questionable methods to extrapolate some orders of magnitude to say that at physical resolution (i.e. solar mass particles) there is no stochasticity in binary evolution. I'm of the opinion this is not the case. The title of this paper is sort of a rebuke to their work.}

\drafting{
    Introduce necessary formulae
}

\begin{equation}\label{eq:quinlan_H}
    \dv{}{t}\left( \frac{1}{a} \right) = H \frac{G\rho}{\sigma}
\end{equation}

\begin{equation}\label{eq:quinlan_K}
    \dv{e}{t} = -K a^{-1} \dv{a}{t}
\end{equation}

\begin{equation}\label{eq:kepler_L}
    L = \mu\sqrt{GMa \left(1-e^2\right)}
\end{equation}

\begin{equation}\label{eq:influence_radius}
    r_\mathrm{infl} = r_\star(<2 M)
\end{equation}

\begin{equation}\label{eq:a_hard}
    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{infl}}{4},
\end{equation}
where $q=M_{\bullet,2}/M_{\bullet,1} < 1$.