### 2023-05-09 papers/paper-eccentricity/sections/discuss.tex

\section{Discussion}

\drafting{
    An interesting point to discuss is the origin of our eccentricity scatter. Early tests of the Gualandris setup seems to indicate that low eccentricity mergers do not demonstrate orbital flips, and are more converged. How to distinguish if the flip is the cause of the eccentricity scatter, or just correlated with eccentricity scatter?
}