### 2024-07-04 papers/paper-corekick/diagrams/dag.tex

\documentclass[border=10pt]{standalone}
\usepackage{newtxtext, newtxmath}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\begin{document}
\begin{tikzpicture}

    %%% define layer 1 %%%
    % density
    \node[obs] (dens) {$\log_{10}\Sigma$};
    % density distribution
    \factor[above=of dens] {dens_D} {right:$\mathcal{N}_\mathrm{likelihood}$} {} {};
    % calculated density
    \node[det, above=of dens_D] (dens_hat) {$\log_{10}\hat{\Sigma}$}; 
    % radius
    \node[obs, above=of dens_hat] (R) {$R$};
    % density error
    \node[latent, right=of dens_hat, yshift=-3em, xshift=-3em] (dens_err) {$\tau_{\log_{10}\Sigma}$};

    % connect
    \factoredge{dens_hat, dens_err} {dens_D} {dens};
    \edge {R} {dens_hat};
    % add the plate
    \plate{level1} {(R)(dens_hat)(dens_err)(dens_D)(dens)} {$\forall\; R$};

    %%% define layer 2 %%%
    % break density
    \node[latent, left=of dens_hat, yshift=10em] (dens_b) {$\log_{10}\Sigma_\mathrm{b}$};
    % inner core slope
    \node[latent, below=of dens_b] (g) {$\gamma$};
    % sersic index
    \node[latent, below=of g] (n) {$n$};
    % transition index
    \node[latent, below=of n] (a) {$\alpha$};
    % core radius
    \node[latent, below=of a] (rb) {$r_\mathrm{b}$};
    % effective radius
    \node[latent, below=of rb] (Re) {$R_\mathrm{e}$};

    % connect
    \edge{dens_b, g, n, a, rb, Re} {dens_hat};
    % add the plate
    \plate{level2} {(dens_b)(g)(n)(a)(rb)(Re)(level1)} {$\forall\; \mathrm{projections}$};

    %%% define layer 3 %%%
    % break radius hypers
    \factor[left=of dens_b, xshift=-0.7em] {dens_b_D} {below:$\mathcal{N}_\mathrm{T[-5,15]}$} {} {};
    \node[hyper, left=of dens_b_D, yshift=2em] (dens_b_mu) {$\mu_{\log_{10}\Sigma_\mathrm{b}}$};
    \node[hyper, below=of dens_b_mu, yshift=2em] (dens_b_err) {$\sigma_{\log_{10}\Sigma_\mathrm{b}}$};
    \factoredge{dens_b_mu, dens_b_err} {dens_b_D} {dens_b};

    % inner slope hypers
    \factor[left=of g, xshift=-1.5em] {g_D} {below:$\mathrm{Exp}_\mathrm{T[0,2]}$} {} {};
    \node[hyper, left=of g_D] (g_lam) {$\lambda_\gamma$};
    \factoredge{g_lam} {g_D} {g};

    % sersic index hypers
    \factor[left=of n, xshift=-1.5em] {n_D} {below:$\mathcal{N}_\mathrm{T[0,20]}$} {} {};
    \node[hyper, left=of n_D, yshift=1em] (n_mu) {$\mu_{n}$};
    \node[hyper, below=of n_mu, yshift=2em] (n_err) {$\sigma_{n}$};
    \factoredge{n_mu, n_err} {n_D} {n};

    % transition index hypers
    \factor[left=of a, xshift=-1.5em] {a_D} {below:$\mathrm{Ray}_\mathrm{T[0,15]}$} {} {};
    \node[hyper, left=of a_D] (a_sig) {$\sigma_\alpha$};
    \factoredge{a_sig} {a_D} {a};

    % core radius hypers
    \factor[left=of rb, xshift=-1.5em] {rb_D} {below:$\mathrm{Ray}_\mathrm{T[0,5]}$} {} {};
    \node[hyper, left=of rb_D] (rb_sig) {$\sigma_{r_\mathrm{b}}$};
    \factoredge{rb_sig} {rb_D} {rb};

    % effective radius hypers
    \factor[left=of Re, xshift=-1.5em] {Re_D} {below:$\mathrm{Ray}_\mathrm{T[0,20]}$} {} {};
    \node[hyper, left=of Re_D] (Re_sig) {$\sigma_{R_\mathrm{e}}$};
    \factoredge{Re_sig} {Re_D} {Re};

    % surface density error hypers
    \factor[above=of dens_err, yshift=4em] {dens_err_D} {left:$\mathcal{N}_\mathrm{T[0,]}$} {} {};
    \node[hyper, above=of dens_err_D, yshift=6em] (dens_err_err) {$\sigma_{\tau_{\log_{10}\Sigma}}$};
    \node[hyper, left=of dens_err_err] (dens_err_mu) {$\mu_{\tau_{\log_{10}\Sigma}}$};
    \factoredge{dens_err_mu, dens_err_err} {dens_err_D} {dens_err};

\end{tikzpicture}
\end{document}