### 2024-10-23 baggins/docs/_build/latex/bagginsbayesiananalysisofgalaxy-galaxyinteractionsinsimulations.tex
Move baggins subdirectory to top level

%% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\ifdefined\pdfpxdimen
   \let\sphinxpxdimen\pdfpxdimen\else\newdimen\sphinxpxdimen
\fi \sphinxpxdimen=.75bp\relax
\ifdefined\pdfimageresolution
    \pdfimageresolution= \numexpr \dimexpr1in\relax/\sphinxpxdimen\relax
\fi
%% let collapsible pdf bookmarks panel have high depth per default
\PassOptionsToPackage{bookmarksdepth=5}{hyperref}

\PassOptionsToPackage{booktabs}{sphinx}
\PassOptionsToPackage{colorrows}{sphinx}

\PassOptionsToPackage{warn}{textcomp}
\usepackage[utf8]{inputenc}
\ifdefined\DeclareUnicodeCharacter
% support both utf8 and utf8x syntaxes
  \ifdefined\DeclareUnicodeCharacterAsOptional
    \def\sphinxDUC#1{\DeclareUnicodeCharacter{"#1}}
  \else
    \let\sphinxDUC\DeclareUnicodeCharacter
  \fi
  \sphinxDUC{00A0}{\nobreakspace}
  \sphinxDUC{2500}{\sphinxunichar{2500}}
  \sphinxDUC{2502}{\sphinxunichar{2502}}
  \sphinxDUC{2514}{\sphinxunichar{2514}}
  \sphinxDUC{251C}{\sphinxunichar{251C}}
  \sphinxDUC{2572}{\textbackslash}
\fi
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amstext}
\usepackage{babel}

\usepackage{tgtermes}
\usepackage{tgheros}
\renewcommand{\ttdefault}{txtt}

\usepackage[Bjarne]{fncychap}
\usepackage{sphinx}

\fvset{fontsize=auto}
\usepackage{geometry}

% Include hyperref last.
\usepackage{hyperref}
% Fix anchor placement for figures with captions.
\usepackage{hypcap}% it must be loaded after hyperref.
% Set up styles of URL: it should be placed after hyperref.
\urlstyle{same}

\addto\captionsenglish{\renewcommand{\contentsname}{Contents:}}

\usepackage{sphinxmessages}
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

\title{BAGGInS: Bayesian Analysis of Galaxy\sphinxhyphen{}Galaxy Interactions in Simulations}
\date{Oct 21, 2024}
\release{1.0}
\author{Alex Rawlings}
\newcommand{\sphinxlogo}{\vbox{}}
\renewcommand{\releasename}{Release}
\makeindex
\begin{document}

\ifdefined\shorthandoff
  \ifnum\catcode`\=\string=\active\shorthandoff{=}\fi
  \ifnum\catcode`\"=\active\shorthandoff{"}\fi
\fi

\pagestyle{empty}
\sphinxmaketitle
\pagestyle{plain}
\sphinxtableofcontents
\pagestyle{normal}
\phantomsection\label{\detokenize{index::doc}}

\sphinxstepscope

\chapter{Python API}
\label{\detokenize{references/API:python-api}}\label{\detokenize{references/API:baggins-api}}\label{\detokenize{references/API::doc}}

\section{BAGGInS}
\label{\detokenize{references/API:baggins}}
\sphinxAtStartPar
Here you can find a list of available functions and classes in the BAGGInS package.

\sphinxstepscope

\subsection{\_backend package}
\label{\detokenize{references/_backend:backend-package}}\label{\detokenize{references/_backend::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/_backend:submodules}}

\subsubsection{\_backend.Logging module}
\label{\detokenize{references/_backend:module-_backend.Logging}}\label{\detokenize{references/_backend:backend-logging-module}}\index{module@\spxentry{module}!\_backend.Logging@\spxentry{\_backend.Logging}}\index{\_backend.Logging@\spxentry{\_backend.Logging}!module@\spxentry{module}}\index{setup\_logger() (in module \_backend.Logging)@\spxentry{setup\_logger()}\spxextra{in module \_backend.Logging}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.Logging.setup_logger}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{\_backend.Logging.}}\sphinxbfcode{\sphinxupquote{setup\_logger}}}
{\sphinxparam{\DUrole{n}{name}}\sphinxparamcomma \sphinxparam{\DUrole{n}{console\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{logfile}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{file\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{capture\_warnings}\DUrole{o}{=}\DUrole{default_value}{True}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Set up a logger instance
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{name}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}) \textendash{} name of logger

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{console\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for console printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{logfile}} (\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} file to print logs to, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{file\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for file printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{capture\_warnings}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} redirect all warnings to the logger, by default True

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{CustomLogger} \textendash{} logger

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{\_backend.States module}
\label{\detokenize{references/_backend:module-_backend.States}}\label{\detokenize{references/_backend:backend-states-module}}\index{module@\spxentry{module}!\_backend.States@\spxentry{\_backend.States}}\index{\_backend.States@\spxentry{\_backend.States}!module@\spxentry{module}}\index{PublishingState (class in \_backend.States)@\spxentry{PublishingState}\spxextra{class in \_backend.States}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.PublishingState}}
\pysigstartsignatures
\pysigline
{\sphinxbfcode{\sphinxupquote{class\DUrole{w}{ }}}\sphinxcode{\sphinxupquote{\_backend.States.}}\sphinxbfcode{\sphinxupquote{PublishingState}}}
\pysigstopsignatures
\sphinxAtStartPar
Bases: \sphinxcode{\sphinxupquote{\_State}}
\index{\_\_init\_\_() (\_backend.States.PublishingState method)@\spxentry{\_\_init\_\_()}\spxextra{\_backend.States.PublishingState method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.PublishingState.__init__}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{\_\_init\_\_}}}
{}
{}
\pysigstopsignatures
\sphinxAtStartPar
Simple object to store the state of a plotting\sphinxhyphen{}style setting

\end{fulllineitems}

\end{fulllineitems}

\index{TmpDirRegister (class in \_backend.States)@\spxentry{TmpDirRegister}\spxextra{class in \_backend.States}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{class\DUrole{w}{ }}}\sphinxcode{\sphinxupquote{\_backend.States.}}\sphinxbfcode{\sphinxupquote{TmpDirRegister}}}
{\sphinxparam{\DUrole{n}{basename}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Bases: \sphinxcode{\sphinxupquote{object}}
\index{\_\_init\_\_() (\_backend.States.TmpDirRegister method)@\spxentry{\_\_init\_\_()}\spxextra{\_backend.States.TmpDirRegister method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.__init__}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{\_\_init\_\_}}}
{\sphinxparam{\DUrole{n}{basename}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Class to hold a register of temporary directories. This means that
temporary directories are particular to the given invocation of
baggins, and can be cleaned up at exit without affecting the
temporary directories of other invocations running concurrently.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{basename}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}) \textendash{} base temporary directory path to which a timestamp will be appended

\end{description}\end{quote}

\end{fulllineitems}

\index{make\_new\_dir() (\_backend.States.TmpDirRegister method)@\spxentry{make\_new\_dir()}\spxextra{\_backend.States.TmpDirRegister method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.make_new_dir}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{make\_new\_dir}}}
{}
{}
\pysigstopsignatures
\sphinxAtStartPar
Make a new temporary directory, and add it to the register

\end{fulllineitems}

\index{register (\_backend.States.TmpDirRegister property)@\spxentry{register}\spxextra{\_backend.States.TmpDirRegister property}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.register}}
\pysigstartsignatures
\pysigline
{\sphinxbfcode{\sphinxupquote{property\DUrole{w}{ }}}\sphinxbfcode{\sphinxupquote{register}}}
\pysigstopsignatures
\end{fulllineitems}

\end{fulllineitems}

\subsubsection{Module contents}
\label{\detokenize{references/_backend:module-_backend}}\label{\detokenize{references/_backend:module-contents}}\index{module@\spxentry{module}!\_backend@\spxentry{\_backend}}\index{\_backend@\spxentry{\_backend}!module@\spxentry{module}}\index{setup\_logger() (in module \_backend)@\spxentry{setup\_logger()}\spxextra{in module \_backend}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.setup_logger}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{\_backend.}}\sphinxbfcode{\sphinxupquote{setup\_logger}}}
{\sphinxparam{\DUrole{n}{name}}\sphinxparamcomma \sphinxparam{\DUrole{n}{console\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{logfile}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{file\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{capture\_warnings}\DUrole{o}{=}\DUrole{default_value}{True}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Set up a logger instance
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{name}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}) \textendash{} name of logger

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{console\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for console printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{logfile}} (\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} file to print logs to, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{file\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for file printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{capture\_warnings}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} redirect all warnings to the logger, by default True

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{CustomLogger} \textendash{} logger

\end{description}\end{quote}

\end{fulllineitems}

\sphinxstepscope

\subsection{analysis package}
\label{\detokenize{references/analysis:analysis-package}}\label{\detokenize{references/analysis::doc}}

\subsubsection{Subpackages}
\label{\detokenize{references/analysis:subpackages}}
\sphinxstepscope

\paragraph{analysis.analysis\_classes package}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-package}}\label{\detokenize{references/analysis.analysis_classes::doc}}

\subparagraph{Submodules}
\label{\detokenize{references/analysis.analysis_classes:submodules}}

\subparagraph{analysis.analysis\_classes.BHBinary module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-bhbinary-module}}

\subparagraph{analysis.analysis\_classes.BHBinaryData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-bhbinarydata-module}}

\subparagraph{analysis.analysis\_classes.Brownian module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-brownian-module}}

\subparagraph{analysis.analysis\_classes.BrownianData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-browniandata-module}}

\subparagraph{analysis.analysis\_classes.ChildSim module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-childsim-module}}

\subparagraph{analysis.analysis\_classes.ChildSimData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-childsimdata-module}}

\subparagraph{analysis.analysis\_classes.CoreKick module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-corekick-module}}

\subparagraph{analysis.analysis\_classes.GaussianProcesses module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-gaussianprocesses-module}}

\subparagraph{analysis.analysis\_classes.GrahamModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-grahammodels-module}}

\subparagraph{analysis.analysis\_classes.HDF5Base module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hdf5base-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesBinary module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiesbinary-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesBinaryData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiesbinarydata-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesSingle module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiessingle-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesSingleData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiessingledata-module}}

\subparagraph{analysis.analysis\_classes.KeplerModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-keplermodels-module}}

\subparagraph{analysis.analysis\_classes.PQModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-pqmodels-module}}

\subparagraph{analysis.analysis\_classes.QuinlanModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-quinlanmodels-module}}

\subparagraph{analysis.analysis\_classes.StanModel module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-stanmodel-module}}

\subparagraph{Module contents}
\label{\detokenize{references/analysis.analysis_classes:module-contents}}

\subsubsection{Submodules}
\label{\detokenize{references/analysis:submodules}}

\subsubsection{analysis.analyse\_ketju module}
\label{\detokenize{references/analysis:analysis-analyse-ketju-module}}

\subsubsection{analysis.analyse\_snap module}
\label{\detokenize{references/analysis:analysis-analyse-snap-module}}

\subsubsection{analysis.general module}
\label{\detokenize{references/analysis:analysis-general-module}}

\subsubsection{analysis.masks module}
\label{\detokenize{references/analysis:module-analysis.masks}}\label{\detokenize{references/analysis:analysis-masks-module}}\index{module@\spxentry{module}!analysis.masks@\spxentry{analysis.masks}}\index{analysis.masks@\spxentry{analysis.masks}!module@\spxentry{module}}\index{get\_all\_id\_masks() (in module analysis.masks)@\spxentry{get\_all\_id\_masks()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_all_id_masks}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_all\_id\_masks}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Return a dict of masks that mask the chosen particles by ID number to a
specific galaxy, assuming each galaxy has a single SMBH in it.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default ‘stars’

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstylestrong{masks} (\sphinxstyleemphasis{dict})

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{keys corresponding to the host galaxy SMBH particle ID, values to pygad.}

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{snapshot.masks.IDMask objects}

\end{itemize}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_all\_radial\_masks() (in module analysis.masks)@\spxentry{get\_all\_radial\_masks()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_all_radial_masks}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_all\_radial\_masks}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{radius}}\sphinxparamcomma \sphinxparam{\DUrole{n}{centre}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_masks}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Return a dict of masks that mask the chosen particles by radial distance to
the desired centre. Either a ball mask or a shell mask may be constructed.
Additionally, the option to initially filter particles by ID to constrain
the mask to those particles belonging to a given galaxy is possible.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{radius}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{tuple}}) \textendash{} radius to constrain the particles to \sphinxhyphen{} can be either a number to
construct a ball mask, or a tuple of (inner\_radius, outer\_radius)
to construct a shell mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{centre}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.UnitArr}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} centre from which the radial measurements should be made, by default
None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID masks to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type to construct the mask for, by default None (all particle
used)

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstylestrong{masks} (\sphinxstyleemphasis{dict}) \textendash{} pygad mask objects, with keys corresponding to the host galaxy SMBH
particle ID number

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} centre method is invalid

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_binding\_energy\_mask() (in module analysis.masks)@\spxentry{get\_binding\_energy\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_binding_energy_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_binding\_energy\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{energy}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_mask}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{None}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Mask particles into bins of binding energy. Assumes that we are in centre of mass coordinates (including velocity): no centring is done.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{energy}} (\sphinxstyleliteralemphasis{\sphinxupquote{array\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{list}}\sphinxstyleliteralemphasis{\sphinxupquote{ of }}\sphinxstyleliteralemphasis{\sphinxupquote{tuples}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} energy bins to mask particles into \sphinxhyphen{}  can either be a regular array, in
which a “ball” mask is created, or a list of tuples, in which a “shell”
mask is created. By default (None) creates a “shell” type mask from the
5\% to 95\% quantile of binding energy in 20 bins evenly spaced in
logscale

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.snapshot.masks.IDMask}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID mask to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{\_type\_}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default None (all)

\end{itemize}

\sphinxlineitem{Yields}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstylestrong{mask} (\sphinxstyleemphasis{pygad.snapshots.mask.IDMask}) \textendash{} mask for that energy interval

\item {} 
\sphinxAtStartPar
\sphinxstylestrong{energy} (\sphinxstyleemphasis{array\sphinxhyphen{}like, list of tuples}) \textendash{} energy sequence

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{str} \textendash{} energy units

\end{itemize}

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} invalid energy input

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_id\_mask() (in module analysis.masks)@\spxentry{get\_id\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_id_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_id\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{bhid}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Obtain a mask that allows filtering of particular particles. Particles are
first filtered to the host galaxy depending on their particle ID number,
where it is necessarily assumed that the particle ID ordering follows:
\begin{quote}

\sphinxAtStartPar
galaxy 1 IDs \textendash{}\textgreater{} galaxy 2 IDs
\end{quote}

\sphinxAtStartPar
To obtain a mask of all particles in a system, the routine can be run
for each particle family, and the masks combined as e.g.:
\begin{quote}

\sphinxAtStartPar
all\_id\_mask = star\_id\_mask \& dm\_id\_mask \& bh\_id\_mask
\end{quote}

\sphinxAtStartPar
\#TODO should this be a separate method??
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{bhid}} (\sphinxstyleliteralemphasis{\sphinxupquote{int}}) \textendash{} SMBH id number we want to find particles around

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default ‘stars’

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{pygad.snapshot.masks.IDMask} \textendash{} object to mask future snapshots of the same simulation

\sphinxlineitem{Raises}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{RuntimeError}} \textendash{} ID ordering is not divided by galaxy

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{RuntimeError}} \textendash{} only galaxy in the system

\end{itemize}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_radial\_mask() (in module analysis.masks)@\spxentry{get\_radial\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_radial_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_radial\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{radius}}\sphinxparamcomma \sphinxparam{\DUrole{n}{centre}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_mask}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{None}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Create a radial\sphinxhyphen{}based mask, can be either a ball or a shell. The default
function arguments are designed for compatability with get\_all\_radial\_masks.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{radius}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{tuple}}) \textendash{} radius to constrain the particles to \sphinxhyphen{} can be either a number to
construct a ball mask, or a tuple of (inner\_radius, outer\_radius)
to construct a shell mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{centre}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.UnitArr}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} centre from which the radial measurements should be made, by default
None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.snapshot.masks.IDMask}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID mask to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type to construct the mask for, by default None (all particle
used)

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstylestrong{mask} (\sphinxstyleemphasis{pygad.snapshot.masks.BallMask}) \textendash{} radial mask of particles

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} radius input is invalid

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{analysis.orbits module}
\label{\detokenize{references/analysis:analysis-orbits-module}}

\subsubsection{analysis.voronoi module}
\label{\detokenize{references/analysis:analysis-voronoi-module}}

\subsubsection{Module contents}
\label{\detokenize{references/analysis:module-contents}}
\sphinxstepscope

\subsection{cosmology package}
\label{\detokenize{references/cosmology:cosmology-package}}\label{\detokenize{references/cosmology::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/cosmology:submodules}}

\subsubsection{cosmology.conversions module}
\label{\detokenize{references/cosmology:module-cosmology.conversions}}\label{\detokenize{references/cosmology:cosmology-conversions-module}}\index{module@\spxentry{module}!cosmology.conversions@\spxentry{cosmology.conversions}}\index{cosmology.conversions@\spxentry{cosmology.conversions}!module@\spxentry{module}}\index{time2z() (in module cosmology.conversions)@\spxentry{time2z()}\spxextra{in module cosmology.conversions}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.conversions.time2z}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.conversions.}}\sphinxbfcode{\sphinxupquote{time2z}}}
{\sphinxparam{\DUrole{n}{t}}\sphinxparamcomma \sphinxparam{\DUrole{n}{H0}\DUrole{o}{=}\DUrole{default_value}{67.36}}\sphinxparamcomma \sphinxparam{\DUrole{n}{pres}\DUrole{o}{=}\DUrole{default_value}{False}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Estimate the redshift for a corresponding cosmic time
from Carmeli 2008
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{t}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} cosmic time since big bang {[}Gyr{]} \sphinxhyphen{}\textgreater{} pres = False
time before present {[}Gyr{]} \sphinxhyphen{}\textgreater{} pres = True

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{H0}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} Hubble constant in km/s/Mpc, by default 100*cosmology{[}‘h’{]}

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{pres}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} t is given from today not Big Bang?, by default False

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} redshift

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{cosmology.cosmology module}
\label{\detokenize{references/cosmology:module-cosmology.cosmology}}\label{\detokenize{references/cosmology:cosmology-cosmology-module}}\index{module@\spxentry{module}!cosmology.cosmology@\spxentry{cosmology.cosmology}}\index{cosmology.cosmology@\spxentry{cosmology.cosmology}!module@\spxentry{module}}\index{angular\_diameter\_distance() (in module cosmology.cosmology)@\spxentry{angular\_diameter\_distance()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.angular_diameter_distance}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{angular\_diameter\_distance}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the angular diameter distance for a flat universe
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} angular diameter distance {[}kpc{]}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_a0r() (in module cosmology.cosmology)@\spxentry{get\_a0r()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.get_a0r}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{get\_a0r}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the value a\_0*r from MBW eq. 3.106, for use in cosmological
distance calculations
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} a0*r {[}kpc{]}

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{AssertionError}} \textendash{} redshift larger than redshift of matter radiation equality

\end{description}\end{quote}

\end{fulllineitems}

\index{luminosity\_distance() (in module cosmology.cosmology)@\spxentry{luminosity\_distance()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.luminosity_distance}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{luminosity\_distance}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the luminosity distance for a flat universe
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} luminosity distance {[}kpc{]}

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{Module contents}
\label{\detokenize{references/cosmology:module-cosmology}}\label{\detokenize{references/cosmology:module-contents}}\index{module@\spxentry{module}!cosmology@\spxentry{cosmology}}\index{cosmology@\spxentry{cosmology}!module@\spxentry{module}}
\sphinxstepscope

\subsection{general package}
\label{\detokenize{references/general:general-package}}\label{\detokenize{references/general::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/general:submodules}}

\subsubsection{general.general module}
\label{\detokenize{references/general:general-general-module}}

\subsubsection{general.pygad\_helper module}
\label{\detokenize{references/general:general-pygad-helper-module}}

\subsubsection{general.units module}
\label{\detokenize{references/general:general-units-module}}

\subsubsection{Module contents}
\label{\detokenize{references/general:module-contents}}
\sphinxstepscope

\subsection{initialise package}
\label{\detokenize{references/initialise:initialise-package}}\label{\detokenize{references/initialise::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/initialise:submodules}}

\subsubsection{initialise.GalaxyIC module}
\label{\detokenize{references/initialise:initialise-galaxyic-module}}

\subsubsection{initialise.MergerIC module}
\label{\detokenize{references/initialise:initialise-mergeric-module}}

\subsubsection{initialise.galaxy\_components module}
\label{\detokenize{references/initialise:initialise-galaxy-components-module}}

\subsubsection{initialise.mergers module}
\label{\detokenize{references/initialise:initialise-mergers-module}}

\subsubsection{Module contents}
\label{\detokenize{references/initialise:module-contents}}
\sphinxstepscope

\subsection{literature package}
\label{\detokenize{references/literature:literature-package}}\label{\detokenize{references/literature::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/literature:submodules}}

\subsubsection{literature.LiteratureTables module}
\label{\detokenize{references/literature:literature-literaturetables-module}}

\subsubsection{literature.bh\_bulge module}
\label{\detokenize{references/literature:literature-bh-bulge-module}}

\subsubsection{literature.density\_profiles module}
\label{\detokenize{references/literature:literature-density-profiles-module}}

\subsubsection{literature.dm\_bulge module}
\label{\detokenize{references/literature:literature-dm-bulge-module}}

\subsubsection{literature.radial\_relations module}
\label{\detokenize{references/literature:literature-radial-relations-module}}

\subsubsection{literature.smbh\_recoil module}
\label{\detokenize{references/literature:literature-smbh-recoil-module}}

\subsubsection{literature.smbh\_spins module}
\label{\detokenize{references/literature:literature-smbh-spins-module}}

\subsubsection{Module contents}
\label{\detokenize{references/literature:module-contents}}
\sphinxstepscope

\subsection{mathematics package}
\label{\detokenize{references/mathematics:mathematics-package}}\label{\detokenize{references/mathematics::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/mathematics:submodules}}

\subsubsection{mathematics.array\_operations module}
\label{\detokenize{references/mathematics:mathematics-array-operations-module}}

\subsubsection{mathematics.coordinates module}
\label{\detokenize{references/mathematics:mathematics-coordinates-module}}

\subsubsection{mathematics.geometry module}
\label{\detokenize{references/mathematics:mathematics-geometry-module}}

\subsubsection{mathematics.statistics module}
\label{\detokenize{references/mathematics:mathematics-statistics-module}}

\subsubsection{Module contents}
\label{\detokenize{references/mathematics:module-contents}}
\sphinxstepscope

\subsection{plotting package}
\label{\detokenize{references/plotting:plotting-package}}\label{\detokenize{references/plotting::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/plotting:submodules}}

\subsubsection{plotting.PlotClasses module}
\label{\detokenize{references/plotting:plotting-plotclasses-module}}

\subsubsection{plotting.custom\_cmaps module}
\label{\detokenize{references/plotting:plotting-custom-cmaps-module}}

\subsubsection{plotting.general module}
\label{\detokenize{references/plotting:plotting-general-module}}

\subsubsection{plotting.plot\_utils module}
\label{\detokenize{references/plotting:plotting-plot-utils-module}}

\subsubsection{plotting.specific\_plots module}
\label{\detokenize{references/plotting:plotting-specific-plots-module}}

\subsubsection{Module contents}
\label{\detokenize{references/plotting:module-contents}}
\sphinxstepscope

\subsection{utils package}
\label{\detokenize{references/utils:utils-package}}\label{\detokenize{references/utils::doc}}

\subsubsection{utils.data\_handling module}
\label{\detokenize{references/utils:utils-data-handling-module}}

\subsubsection{utils.dataset\_operations module}
\label{\detokenize{references/utils:utils-dataset-operations-module}}

\subsubsection{utils.get\_cl\_args module}
\label{\detokenize{references/utils:utils-get-cl-args-module}}

\subsubsection{utils.os\_operations module}
\label{\detokenize{references/utils:utils-os-operations-module}}

\subsubsection{utils.param\_io module}
\label{\detokenize{references/utils:utils-param-io-module}}

\subsubsection{Module contents}
\label{\detokenize{references/utils:module-contents}}
\sphinxstepscope

\subsection{visualisation package}
\label{\detokenize{references/visualisation:visualisation-package}}\label{\detokenize{references/visualisation::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/visualisation:submodules}}

\subsubsection{visualisation.animation module}
\label{\detokenize{references/visualisation:visualisation-animation-module}}

\subsubsection{Module contents}
\label{\detokenize{references/visualisation:module-contents}}

\renewcommand{\indexname}{Python Module Index}
\begin{sphinxtheindex}
\let\bigletter\sphinxstyleindexlettergroup
\bigletter{\_}
\item\relax\sphinxstyleindexentry{\_backend}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend}}
\item\relax\sphinxstyleindexentry{\_backend.Logging}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend.Logging}}
\item\relax\sphinxstyleindexentry{\_backend.States}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend.States}}
\indexspace
\bigletter{a}
\item\relax\sphinxstyleindexentry{analysis.masks}\sphinxstyleindexpageref{references/analysis:\detokenize{module-analysis.masks}}
\indexspace
\bigletter{c}
\item\relax\sphinxstyleindexentry{cosmology}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology}}
\item\relax\sphinxstyleindexentry{cosmology.conversions}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology.conversions}}
\item\relax\sphinxstyleindexentry{cosmology.cosmology}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology.cosmology}}
\end{sphinxtheindex}

\renewcommand{\indexname}{Index}
\printindex
\end{document}